---
title: "ETC5521 Assignment 1"
subtitle: ""
team: Quokka
author:
  - Dea Avega Editya
  - Siyi Li
date: "`r Sys.Date()`"
output: 
  html_document
---

[This assignment is for ETC5521 Assignment 1 by Team `r rmarkdown::metadata$team` comprising of `r knitr::combine_words(rmarkdown::metadata$author)`.]{style="color:#006DAE;"}

```{r read-data}
library(tidyverse)
library(janitor)
library(readr)

```


```{r read-data}
# Get the Data

blackpast <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-16/blackpast.csv')
census <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-16/census.csv')
slave_routes <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-16/slave_routes.csv')
african_names <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-16/african_names.csv')

```


# Introduction and motivation

We are motivated to explore dark history of slavery in United States of America and how this moment changed over time of hundred years. Through the data we will look at how the situation at the moment was and what can we learn from the history that maybe has correlation with current racist behavior toward black people especially in the USA.

[FILL] Give the bigger picture of the data, and inspire the reader to learn more about the problem by reading your analysis. 

# Data description

<!-- Write your data description under the header “Data description” in index.Rmd. A data description should tell the reader the source of the data, and detail the structure of the data, including variables, time frame of collection, collection methods. -->

Our datasets are retrieved from github repository of tidytuesday project (https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-06-16/readme.md) which has original source from US Census's Archives, Slave Voyages, and Black Past. 

The data consists of 4 datasets as follows:
1. Census (in csv format)
  The dataset record the total slave populations across the United States during the slavery era and has 8 variables (region, division , year, total, white, black, black_free and black_slave) and 102 observation. The data is collected from a historical US census data with time period from 1790 to 1870 

2. Slave_routes (in csv format) 
  the data covers transportation of slaves and has 8 variables (voyage_id, ship_name, port_origin, place_of_purchase, port_arrival, year_arrival, n_slaves_arrived, and captains_name) and 31.660 observations covers some details in slave transportation, from their origin places, trade hub, up to their arrival places. The record is from 1514 to 1866.
  
3. African_names (in csv format)
  The dataset contains a history of freed slaves with 11 variables (id, voyage id , name, gender, age, height, ship name, year arrival, port embark, and country origin) and 91.490 observations. The record is from 1808 to 1862.

4. Blackpast (in csv format)
  The dataset covers details around the slow granting of various rights to African-Americans, post-slavery brutality, violence and racism, as well as celebrations of achievements across many different categories. It has 6 variables (year, event, subject, country, state and era) and 896 observations. The record is from 1492 to 2009

Using these above datasets, we will explore several main questions like:
1. What is the demographic of slaves?
2. Which region of the USA that most exploit the practice?   
3. How is the proportion of white and black slaves, as well as black slaves and black free?


# Analysis and findings

<!-- Brainstorm what questions would be possible to answer with this data (and only this data). Articulate the primary question, and then three secondary questions. -->

[FILL] Should include at least one plot or numerical summary for each of your questions, that helps the reader arrive at an answer. You should also write paragraphs describing the methods, summaries and findings. 

# References
